---
title: "TurningAnalysis"
format: pdf
editor: source
---

# Importing

## Library import and global variables
```{r libraries and functions, message = FALSE, output = FALSE}
library(here)
here::i_am('Code/TurningAnalysis.qmd')
source(here('Code/Functions/CalcFunctions.R'))
library(easystats)

library(conflicted)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::lag)

importLibraries()

library(GGally)
library(multcomp)

bodypartorder <- c('Snout','GillCenterLow','PelvFinsCenter','MidPelvAnal',
                   'AnalBase','MidAnalBasePeduncle','Peduncle')
dt = 1/60
```

## Read in data and select relevant columns
```{r data import, message=FALSE}
maindata <- read_csv(here("ProcessedData/fulldata.csv")) |> 
  mutate(bodypart = factor(bodypart, levels = bodypartorder),
         ID = factor(ID),
         Individual = factor(Individual)) |> 
  select(c('ID', 'Speed.BLs', 'Speed.group', 'frame', 't.sec',
           'bodypart', 'xs.cm', 'ys.cm', 'Individual', 'Trial',
           'Turn', 'FullBL', 'mass.g', 's0'))

head(maindata)
```

```{r}
maindata |> 
  distinct(Individual)
```

```{r}
maindata |> 
  distinct(Individual, Trial, Turn) |> 
  group_by(Individual, Trial) |> 
  summarize(maxturnnum = max(Turn),
            maxdone = if_else(maxturnnum < 16, 16, maxturnnum),
            nexcluded = length(setdiff(seq(1,maxdone), Turn)))
```


# Snout Velocity

## We can look at the velocity of the snout in greater detail

We do this by calculating the change in angle, then the change in the change in angle over time

```{r look at snout velocity, message = FALSE}
SnoutAngChangeButter <- signal::butter(3, 4/(60/2), 'low')

SnoutVelocityDataPreTrim <- maindata |> 
  calcAngChange() |> 
  runSnoutAngButter(SnoutAngChangeButter) |> 
  calcAngVel(dt) |>
  AddBackSpeed(maindata) 

SnoutVelocityData <- SnoutVelocityDataPreTrim |> trimFrames()
```

## And we can compare snout angular velocity, first with plots
```{r look at how long the turns take, message = FALSE, warning = FALSE}
snout_angle_plot <- SnoutVelocityData |> SnoutAnglePlot()

meansnoutvelslow <- SnoutVelocityData |> calcMeanSnoutVelocity('Slow', 1)

snout_ang_veL_slow_plot <- 
  SnoutVelocityData |> SlowSnoutVelocityPlot(meansnoutvelslow)

meansnoutvelfast <- SnoutVelocityData |> calcMeanSnoutVelocity('Fast', .75)

snout_ang_veL_fast_plot <- 
  SnoutVelocityData |> FastSnoutVelocityPlot(meansnoutvelfast)

snout_angle_plot / (snout_ang_veL_fast_plot | snout_ang_veL_slow_plot) +
  plot_annotation(tag_levels = 'A')

ggsave('SnoutAngularVelocity.pdf', path = here("ProcessedData/Figures/"), width = 6, height = 4, units = 'in')
```

### Then with stats
```{r}
MeanVelStats <- SnoutVelocityData |> 
  summarize(ID = first(ID),
            Speed.group = first(Speed.group),
            meansnoutvel.degps = mean(snoutangle.degps, na.rm = TRUE)) |> 
  mutate(Individual = substr(ID, 1, 4))

MaxVelStats <- SnoutVelocityData |> 
  summarize(ID = first(ID),
            Speed.group = first(Speed.group),
            maxsnoutvel.degps = max(snoutangle.degps, na.rm = TRUE)) |> 
  mutate(Individual = substr(ID, 1, 4))

meansnoutvelmodel <- lmer(meansnoutvel.degps ~ Speed.group + (1 | Individual), data = MeanVelStats)
Anova(meansnoutvelmodel)

maxsnoutvelmodel <- lmer(maxsnoutvel.degps ~ Speed.group + (1 | Individual), data = MaxVelStats)
Anova(maxsnoutvelmodel)

MeanVelStatsSlow <- MeanVelStats |> filter(Speed.group == 'Slow')
MeanVelStatsFast <- MeanVelStats |> filter(Speed.group == 'Fast')
summary(MeanVelStatsSlow$meansnoutvel.degps)
sd(MeanVelStatsSlow$meansnoutvel.degps, na.rm = TRUE)
summary(MeanVelStatsFast$meansnoutvel.degps)
sd(MeanVelStatsFast$meansnoutvel.degps, na.rm = TRUE)

MaxVelStatsSlow <- MaxVelStats |> filter(Speed.group == 'Slow')
MaxVelStatsFast <- MaxVelStats |> filter(Speed.group == 'Fast')
summary(MaxVelStatsSlow$maxsnoutvel.degps)
sd(MaxVelStatsSlow$maxsnoutvel.degps, na.rm = TRUE)
summary(MaxVelStatsFast$maxsnoutvel.degps)
sd(MaxVelStatsFast$maxsnoutvel.degps, na.rm = TRUE)
```

```{r}
check_model(meansnoutvelmodel)
```

```{r}
check_model(maxsnoutvelmodel)
```

# Moment of Inertia

Moment of inertia is $$ I = \sum m_i r_i^2 $$

And can integrate MoI to get a sense of how bent the body is overall during the turn

## First, let's add each individual's maximum theoretical MoI
```{r}
MaxMoI <- maindata |> calcMaxMoI()
```

## Now we calculate the MoI in each frame
```{r calculate MoI and look at its integral, message = FALSE}
MoIPreTrim <- maindata |> 
  calcMoI() |> 
  left_join(MaxMoI, by = c('Individual'))

MoI <- MoIPreTrim |> trimFrames()

MoI <- SnoutVelocityData |> select(ID, frame, snoutangle.degps) |> right_join(MoI, by = c('ID', 'frame'))

```

## Figure with body traces and MoI

### How do the traces for a fast vs. a slow turn (respectively) look?

```{r}
set_centered_range <- function(rng, width) {
  ctr <- (rng[1] + rng[2]) / 2
  rng <- ctr + c(-width/2, width/2)
  
  return(rng)
}
```

```{r body traces for a fast vs. slow turn, warning = FALSE}
fasttrace <- maindata %>% 
  bodyTracesPlot('BG63_14_8', 4, reverse_x = TRUE) + 
  scale_color_gradient(low = "#ffeddd", high = "firebrick") +
  scale_x_continuous(limits = \(x) set_centered_range(range(x), 20))

fasttrace

slowtrace <- maindata %>% 
  bodyTracesPlot('BG63_17_5', 4, reverse_x = TRUE) + 
  scale_color_gradient(low = "#d9f1f1", high = "deepskyblue") +
  scale_x_continuous(limits = \(x) set_centered_range(range(x), 20))

slowtrace
```


```{r}
meanmoislow <- MoI |> calcMeanMoI('Slow',1)
slow_moi_plot <- MoI |> SlowMoIPlot(meanmoislow)

meanmoifast <- MoI |> calcMeanMoI('Fast',.75)

fast_moi_plot <- MoI |> FastMoIPlot(meanmoifast)
```

Show the combined figure
```{r}
(slowtrace | slow_moi_plot) / (fasttrace | fast_moi_plot)
ggsave('MoI.pdf', path = here("ProcessedData/Figures/"), width = 6.5, height = 4)
```

### And run some stats
```{r}
MoIpercstats <- MoI |> 
  group_by(ID) |> 
  mutate(percMoI = MoI.gcm2/MaxMoI.gcm2) |> 
  group_by(ID) |> 
  summarize(ID = first(ID),
            Individual = first(Individual),
            Speed.group = first(Speed.group),
            minpercMoI = min(percMoI),
            meanpercMoI = mean(percMoI))

MoIslow <- MoIpercstats |> 
  filter(Speed.group == 'Slow')

MoIfast <- MoIpercstats |> 
  filter(Speed.group == 'Fast')

summary(MoIslow$minpercMoI)
sd(MoIslow$minpercMoI)

summary(MoIfast$minpercMoI)
sd(MoIfast$minpercMoI)

minpercMoImodel <- lmer(minpercMoI ~ Speed.group + (1 | Individual), data = MoIpercstats)
Anova(minpercMoImodel)

summary(MoIslow$meanpercMoI)
sd(MoIslow$meanpercMoI)

summary(MoIfast$meanpercMoI)
sd(MoIfast$meanpercMoI)

meanpercMoImodel <- lmer(meanpercMoI ~ Speed.group + (1 | Individual), data = MoIpercstats)
Anova(meanpercMoImodel)
```

```{r}
check_model(minpercMoImodel)
```

```{r}
check_model(meanpercMoImodel)
```

## Now we compare MoI and speed with stats and plots
```{r}
MeanMoIMeanSnoutAng <- SnoutVelocityData |> 
  group_by(ID) |> 
  calcMeanMoIMeanSnoutVel() 

meanSnoutAng.lm <- lm(meansnoutangle.degps ~ meanpercMoI, data = MeanMoIMeanSnoutAng)
summary(meanSnoutAng.lm)

MoISnoutAngFast <- MeanMoIMeanSnoutAng |> AddBackSpeed(maindata) |> filter(Speed.group == 'Fast')
meanFastSnoutAng.lm <- lm(meansnoutangle.degps ~ meanpercMoI, data = MoISnoutAngFast)
summary(meanFastSnoutAng.lm)

MoISnoutAngSlow <- MeanMoIMeanSnoutAng |> AddBackSpeed(maindata) |> filter(Speed.group == 'Slow')
meanSlowSnoutAng.lm <- lm(meansnoutangle.degps ~ meanpercMoI, data = MoISnoutAngSlow)
summary(meanSlowSnoutAng.lm)

MeanMoIMeanSnoutAng |> MeanMoiMeanSnoutAngVelPlot(meanSnoutAng.lm, meanFastSnoutAng.lm, meanSlowSnoutAng.lm) +
  theme(legend.position = 'none')

ggsave('SnoutVelMoI.pdf', path = here("ProcessedData/Figures/"), width = 3.5, height = 2.5, units = 'in', dpi = 300)
```

# Torque

Torque is the $$ \tau = \frac{d}{dt} \sum I_i \omega_i $$ If we work out the derivative, we end up with $$ \tau = \sum m_i \left(2 r_i \omega_i \dot{r_i} + r_i^2 \dot{\omega_i} \right) $$

## Calculate torque for the whole dataset
```{r whole dataset torque calcuation, message = FALSE}
torquedataPreTrim <- maindata |> calcTorque()

torquedata <- torquedataPreTrim |> trimFrames()
```

# Timing

## First we look at timing of snout angular velocity and MoI
```{r}
MoISnoutVelTiming <- MoI |> 
  mutate(maxT.sec = max(t.sec, na.rm = TRUE)) |> 
  na.omit() |> 
  reframe(maxT.sec = first(maxT.sec),
          tmaxVel = first(t.sec[which.max(snoutangle.degps)]),
          tminMoI = first(t.sec[which.min(MoI.gcm2)]),
          Speed.group = first(as.character(Speed.group))) 

MoISnoutVelTiming |> 
  MoISnoutVelTimingPlot() +
  theme(legend.position = 'none')

ggsave('VelocityMoITiming.pdf', path = here("ProcessedData/Figures/"), width = 3.5, height = 2.5, units = 'in', dpi = 300)

MoIvelTiming <- MoISnoutVelTiming %>% 
  group_by(ID) %>% 
  pivot_longer(cols = c(tmaxVel, tminMoI),
               names_to = 'var',
               values_to = 'val') |> 
  reframe(ID = first(ID),
          MoIVelRatio = lead(val)/val) |> 
  na.omit() |> 
  mutate(MoIorVel = case_when(MoIVelRatio > 1 ~ 'MoISecond',
                              MoIVelRatio == 1 ~ 'SameTime',
                              MoIVelRatio < 1 ~ 'MoIFirst'))
```

### And run the stats
```{r}
lmsnoutmoi <- MeanMoIMeanSnoutAng |> AddBackSpeed(maindata)

MeanMoIMeanSnoutAngStats <- MeanMoIMeanSnoutAng |> AddBackSpeed(maindata)
TotalSnoutAng.lm <- lm(meansnoutangle.degps ~ meanpercMoI*Speed.group, data = MeanMoIMeanSnoutAngStats)
summary(TotalSnoutAng.lm)

Anova(lmer(meansnoutangle.degps ~ meanpercMoI + (1 | Individual), data = MoISnoutAngSlow))
Anova(lmer(meansnoutangle.degps ~ meanpercMoI + (1 | Individual), data = MoISnoutAngFast))

meansnoutanglemodel <- lmer(meansnoutangle.degps ~ meanpercMoI*Speed.group + (1 | Individual), data = MeanMoIMeanSnoutAngStats)
Anova(meansnoutanglemodel)
```

```{r check_meansnoutanglemodel}
check_model(meansnoutanglemodel)
```

## Then we look into the timing of MoI minimization and torque maximization
```{r torquemoitimingplot}
TurnTimeData <- torquedata |> 
  select(ID, frame, torque.gcm2ps2, t.sec) |> 
  left_join(MoI, by = c('ID', 'frame', 't.sec')) 

moitorqueplot1 <- TurnTimeData |> TorqueMoITimingPlot() +
  coord_fixed() +
  theme(legend.position = "none")
moitorqueplot1
```

### And run the stats
```{r torquemoitimingstats}
torquetimingstats <- TurnTimeData |> 
  group_by(ID) |> 
  reframe(tmaxT = t.sec[which.max(torque.gcm2ps2)],
          tminMoI = t.sec[which.min(MoI.gcm2)],
          Speed.group = as.character(Speed.group)) %>% 
  mutate(torquebeforeMoI = tmaxT < tminMoI) |> 
  group_by(ID) |> 
  summarize(ID = first(ID),
            Speed.group = first(Speed.group),
            torquebeforeMoI = first(torquebeforeMoI))

torquetimingstatsslow <- torquetimingstats |> filter(Speed.group == 'Slow')
torquetimingstatsfast <- torquetimingstats |> filter(Speed.group == 'Fast')

summary(torquetimingstatsslow$torquebeforeMoI)
summary(torquetimingstatsfast$torquebeforeMoI)

TurnTimeData.stats <- TurnTimeData |> 
  group_by(ID) |> 
  reframe(tmaxT = t.sec[which.max(torque.gcm2ps2)],
          tminMoI = t.sec[which.min(MoI.gcm2)],
          Speed.group = as.character(first(Speed.group))) |> 
  group_by(ID) |>
  mutate(TimeDiff.sec = tminMoI - tmaxT) |>
  mutate(Individual = substr(ID,1,4)) |> 
  group_by(Individual) |> 
  mutate(IndividualNum = cur_group_id())

turntimeslow <- TurnTimeData.stats |> filter(Speed.group == 'Slow')
turntimefast <- TurnTimeData.stats |> filter(Speed.group == 'Fast')

summary(turntimeslow$TimeDiff.sec)
sd(turntimeslow$TimeDiff.sec)
summary(turntimefast$TimeDiff.sec)
sd(turntimefast$TimeDiff.sec)

torque_moi_timing_model <- lmer(TimeDiff.sec ~ Speed.group + (1 | IndividualNum), data = TurnTimeData.stats)
Anova(torque_moi_timing_model)
```

```{r checktorquemoitimingmodel}
check_model(torque_moi_timing_model)
```

Power analysis

May need to install the "mixedpower" package from GitHub
```
if (!require("devtools")) {
    install.packages("devtools", dependencies = TRUE)
}
devtools::install_github("DejanDraschkow/mixedpower") # mixedpower is hosted on GitHub
```

```{r torquemoi_power_analysis}
#| eval: false
# this block takes a while to run, so don't do it every time we render

library(mixedpower)

power_torque_moi <- mixedpower(torque_moi_timing_model, 
                               data = TurnTimeData.stats,
                               fixed_effects = c('Speed.group'),
                               simvar = 'IndividualNum',
                               steps = c(3,4,5,6,7,8,10,15,20),
                               critical_value = 5, n_sim = 100)
```

## Then we look at the timing differences in more detail
```{r torquemoitimingplot2}
TurnTimeData2 <- TurnTimeData |> 
  group_by(ID) |> 
  reframe(tmaxT = t.sec[which.max(torque.gcm2ps2)],
          tminMoI = t.sec[which.min(MoI.gcm2)],
          maxangvel.degps = max(snoutangle.degps, na.rm = TRUE),
          Speed.group = as.character(first(Speed.group))) |> 
  group_by(ID) |>
  mutate(TimeDiff.sec = tmaxT - tminMoI) |>
  mutate(Individual = substr(ID,1,4))

moitorqueplot2 <- TurnTimeData2 |> TorqueMoITimingPlot2()
moitorqueplot2
```



```{r}
angvel_timediff_model <- lmer(maxangvel.degps ~ TimeDiff.sec + (1 | Individual), data = TurnTimeData2)
Anova(angvel_timediff_model)

```
```{r}
angvel_timediff_plot <-
  plot(estimate_expectation(angvel_timediff_model, data = "grid")) +
  # ggplot(aes(x = TimeDiff.sec, y = estimate,
  #            ymin = conf.low, ymax = conf.high)) +
  # geom_ribbon(alpha = 0.2) +
  # geom_line() +
  geom_point(data = TurnTimeData2, 
             inherit.aes = FALSE,
             aes(x = TimeDiff.sec, y = maxangvel.degps, 
                 color = Speed.group, shape = Speed.group)) +
  scale_color_manual(values = c('firebrick', 'deepskyblue')) +
    scale_shape_manual(values = c(2, 19)) +
  theme_classic() +
  theme(
    legend.position = 'none',
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 8, face = "bold"),
    legend.text = element_text(size = 8),
    legend.key.size = unit(.25, "cm"),
    plot.margin = grid::unit(c(.05, .05, .05, .05), "cm")
  ) +
  labs(color = 'Turn Speed') +
  xlab('Time difference (s)') +
  ylab('Maximum angular\nvelocity (deg s-1)') +
  xlim(-1,1)
  
angvel_timediff_plot
```


```{r}
plot(estimate_grouplevel(angvel_timediff_model))
```

```{r}
(moitorqueplot1 | moitorqueplot2 | angvel_timediff_plot) +
  plot_layout(widths = 1) +
  plot_annotation(tag_levels = 'A')
ggsave(here('ProcessedData','Figures','MoiTorqueTiming.pdf'), width = 6.5, height = 2.2, units = 'in')
```


### And run the stats to compare the time differences to 0
```{r zerotimediffstats}
Anova(lmer(TimeDiff.sec ~ Speed.group + (1 | Individual), data = TurnTimeData2))

linmod <- lmer(TimeDiff.sec ~ Speed.group + (1 | Individual), data = TurnTimeData2)
Anova(linmod)
summary(linmod)

# Get estimated means for each group
group_means <- emmeans(linmod, ~ Speed.group)

# Test if each group's mean is different from zero
contrast(group_means, method = "identity")
```

```{r checklinmod}
check_model(linmod)
```

# Snout and Tail

## First we calculate snout and tail angular velocity
```{r snouttailangvel, warning = FALSE}
TailAngChangeButter <- signal::butter(3, 4/(60/2), 'low')
taildataPreTrim <- maindata |> 
  calcAngChangeTail() |> 
  runTailAngButter(TailAngChangeButter) |> 
  calcAngVelTail(dt)|> 
  AddBackSpeed(maindata)

taildata <- taildataPreTrim |> trimFrames()

reducedtaildata <- taildata |> 
  select(ID, t.sec, tailangle.degps, Speed.group)

reducedSnoutVelocityData <- SnoutVelocityData |> 
  select(ID, t.sec, snoutangle.degps)

SnoutTailAngularVel <- reducedtaildata |> left_join(reducedSnoutVelocityData)
```

## Then we plot a representative trial
```{r snouttailangvelplot, warning = FALSE}
snouttailangvel_plot <-
  SnoutTailAngularVel |> filter(ID == 'BG63_14_10') |> 
  select(c(ID, t.sec, tailangle.degps, snoutangle.degps)) |> 
  mutate(tsnoutmax = t.sec[which.max(snoutangle.degps)],
         ttailmax = t.sec[which.max(tailangle.degps)]) |> 
  glimpse() |> 
  pivot_longer(cols = c(-ID, -t.sec, -tsnoutmax, -ttailmax))|> 
  SnoutVsTailAngularVelocityPlot() +
  theme(legend.position = 'none')

snouttailangvel_plot
```

```{r}
tsnoutmax <- 0.516666666666
ttailmax <- 0.68333333333

maindata |> 
  filter(ID == 'BG63_14_10') |> 
  filter(near(t.sec, tsnoutmax, tol = 1/60) |
           near(t.sec, ttailmax, tol = 1/60)) |> 
  distinct(frame)

angvelmidlines_plot <- 
  maindata |> 
  filter(ID == 'BG63_14_10') |> 
  filter(near(t.sec, tsnoutmax, tol = 1/60) |
           near(t.sec, ttailmax, tol = 1/60)) |> 
  ggplot(aes(xs.cm, ys.cm, color = factor(frame))) +
  geom_path(aes(group = frame)) +
  geom_point(data = ~ filter(.x, 
                             bodypart == "Snout")) +
  coord_fixed() +
  scale_x_reverse() +
  theme_void() +
  theme(legend.position = "none")  

angvelmidlines_plot

```


## And we look at mins and maxes

First we take only the highest snout and tail velocities for each trial
```{r snouttailsummary}
SnoutTailSummary <- SnoutTailAngularVel |> 
  group_by(ID) |> 
  filter(snoutangle.degps == max(snoutangle.degps, na.rm = TRUE) | tailangle.degps == max(tailangle.degps, na.rm = TRUE)) |> 
  mutate(Max = case_when(snoutangle.degps == max(snoutangle.degps, na.rm = TRUE) ~ 'Max Snout',
                         tailangle.degps == max(tailangle.degps, na.rm = TRUE) ~ 'Max Tail'))
```

## Now we plot the maxes

```{r snouttaildotplot, warning = FALSE, message = FALSE}
SnoutTailDotplot <- SnoutTailSummary |> 
  group_by(ID, t.sec) |> 
  summarise(ID = first(ID),
            Speed.group = first(Speed.group),
            SnoutVel.degps = first(snoutangle.degps),
            TailVel.degps = first(tailangle.degps),
            Max = first(Max)) |> 
  pivot_longer(cols = ends_with('.degps'),
               names_to = 'Bodypart',
               values_to = 'AngularVelocity')

SnoutvTailAngVel_plot <- SnoutTailDotplot |> SnoutTailComp('Max Snout')
SnoutvTailAngVel_plot

TailvSnoutAngVel_plot <- SnoutTailDotplot |> SnoutTailComp('Max Tail')
TailvSnoutAngVel_plot
```

```{r}
snouttailangvel_plot + 
  SnoutvTailAngVel_plot + 
  TailvSnoutAngVel_plot +
  plot_layout(nrow = 3) +
  plot_annotation(tag_levels = 'A')

ggsave(here('ProcessedData', 'Figures', 'SnoutTailAngleVel.pdf'),
       width = 2.5, height = 4, units = 'in')
```

### And run the stats
```{r snouttailstats}
snoutfastest <- SnoutTailSummary |> filter(Max == 'Max Snout')
summary(snoutfastest$snoutangle.degps)
sd(snoutfastest$snoutangle.degps)
summary(snoutfastest$tailangle.degps)
sd(snoutfastest$tailangle.degps)

snoutfasteststats <- snoutfastest |> 
  pivot_longer(cols = ends_with('.degps'),
               names_to = 'Bodypart',
               values_to = 'AngularVelocity') |> 
  mutate(Individual = substr(ID,1,4))
Anova(lmer(AngularVelocity ~ Bodypart + (1 | Individual), data = snoutfasteststats))

tailfastest <- SnoutTailSummary |> filter(Max == 'Max Tail')
summary(tailfastest$snoutangle.degps)
sd(tailfastest$snoutangle.degps)
summary(tailfastest$tailangle.degps)
sd(tailfastest$tailangle.degps)

tailfasteststats <- tailfastest |> 
  pivot_longer(cols = ends_with('.degps'),
               names_to = 'Bodypart',
               values_to = 'AngularVelocity') |> 
  mutate(Individual = substr(ID,1,4))
Anova(lmer(AngularVelocity ~ Bodypart + (1 | Individual), data = tailfasteststats))
```

# Pectoral Fins

## Import data
```{r}
pecfindatapaths <- list.files(here("RawData/PecFins"), pattern = "[.]csv$", full.names = TRUE)
pecfindata <- pecfindatapaths |> map(read_csv_filename) |> list_rbind()

pecfinturndetails <- readxl::read_excel(here("RawData/PecFins/PecFinTurnFrames.xlsx"))
```

## Format data
```{r}
pecfindata <- pecfindata |> pecfinformat()
```

## Now we start analysis by counting number of pec fin strokes
```{r}
PecStrokeCount <- pecfindata |> calcPecStrokeNumber(maindata)
```

## And we calculate pec fin strokes per unit time
```{r}
PecStrokeCountTime <- pecfinturndetails |> 
  rowwise() |> 
  mutate(totalframes = EndFrame - StartFrame,
         totaltime = totalframes*dt) |> 
  select(c(ID, totaltime)) |> 
  right_join(PecStrokeCount, by = 'ID') |> 
  group_by(ID) |> 
  mutate(StrokesPerTime.sps = TotalPecStrokes/totaltime,
         Individual= substr(ID,1,4))
```

## How about looking at full time of turn with plots and stats?
```{r}
pecstrokes <- PecStrokeCountTime |> select(ID, totaltime, StrokesPerTime.sps) |> right_join(PecStrokeCount, by = 'ID')

totalpecstrokes.lm <- lm(TotalPecStrokes ~ totaltime, data = pecstrokes)
summary(totalpecstrokes.lm)
totalpecstrokesmodel <- lmer(TotalPecStrokes ~ Speed.group + (1 | Individual), data = pecstrokes)
Anova(totalpecstrokesmodel)

pecstroke_fig <-
  pecstrokes |> TotalPecStrokesPlot(totalpecstrokes.lm)

strokespertimemodel <- lmer(StrokesPerTime.sps ~ Speed.group + (1 | Individual), data = pecstrokes)
Anova(strokespertimemodel)

pecfreq_fig <-
  pecstrokes |> PecStrokesPerTimePlot(pecstrokes)

pecstroke_fig + pecfreq_fig
```

```{r}
check_model(totalpecstrokesmodel)
```

```{r}
check_model(strokespertimemodel)
```

## Pec fin angles calculations
```{r}
pecvectorangles <- pecfindata |> calcPecFinAngles()
```

## Now we look for a role for fin side with plots
```{r}
pecfinamp_fig <-
  pecvectorangles |> PecFinAmplitudePlot()

pecfinamp_fig
```

### And with stats
```{r}
slowpecvec <- pecvectorangles |> filter(Speed.group == 'Slow')
Anova(lmer(pecangle.deg ~ side + (1 | Individual), data = slowpecvec))

fastpecvec <- pecvectorangles |> filter(Speed.group == 'Fast')
Anova(lmer(pecangle.deg ~ side + (1 | Individual), data = fastpecvec))

insidepecvec <- pecvectorangles |> filter(side == 'Inside')
Anova(lmer(pecangle.deg ~ Speed.group + (1 | Individual), data = insidepecvec))

outsidepecvec <- pecvectorangles |> filter(side == 'Outside')
Anova(lmer(pecangle.deg ~ Speed.group + (1 | Individual), data = outsidepecvec))

weirdpecvec1 <- pecvectorangles |> filter(side == 'Inside' & Speed.group == 'Slow')
weirdpecvec2 <- pecvectorangles |> filter(side == 'Outside' & Speed.group == 'Fast')
weirdpecvec <- rbind(weirdpecvec1, weirdpecvec2)
Anova(lmer(pecangle.deg ~ Speed.group + (1 | Individual), data = weirdpecvec))
```

## What about backing?
```{r}
backingdata <- pecvectorangles |> calcNumOfBackingStrokes(90)

backingdata <- PecStrokeCountTime |> 
  select(ID,totaltime) |> 
  right_join(backingdata, by = 'ID')
```

## Now we plot backing data against time
```{r}
backingdata.lm <- lm(numberofbackingstrokes ~ totaltime, data = backingdata)
summary(backingdata.lm)

backing_fig <-
  backingdata |> 
  BackingDataPlot(backingdata.lm)

backing_fig
```

### And do the stats
```{r}
backingdata |> 
  filter(Speed.group == 'Slow') |> 
  summary()

backingdata |> 
  filter(Speed.group == 'Fast') |> 
  summary()

Anova(lmer(numberofbackingstrokes ~ Speed.group + (1 | Individual), data = backingdata))
```

Put together the figure.

```{r}
pecstroke_fig + pecfreq_fig + pecfinamp_fig + backing_fig &
  plot_annotation(tag_levels = 'A')

ggsave(here('ProcessedData', 'Figures', 'PectoralFins.pdf'),
       width=6, height=4, units='in')
```

# Momentum

## We estimate momentum
```{r}
MoIsummary <- MoIPreTrim |> 
  select(ID, frame, MoI.gcm2, MaxMoI.gcm2)

angularmomentumdata <- SnoutVelocityDataPreTrim |> calcAngularMomentum(MoIsummary)
linearmomentumdata <- maindata |> prepLinearMomentum()
momentumdata <- linearmomentumdata |> right_join(angularmomentumdata, by = c('ID', 'frame'))

momentumdata <- maindata |>
  group_by(ID) |> 
  summarise(FullBL = first(FullBL)) |> 
  right_join(momentumdata, by = 'ID')

normalizedmomentumdata <- momentumdata |> calcLinearMomentum(maindata) |> normalizeMomentum()

turnmomentumdata <- normalizedmomentumdata |> 
  trimFrames() |> 
  group_by(ID) |> 
  mutate(maxframe = max(frame, na.rm = TRUE),
    # Time = case_when(frame <= maxframe/2 ~ '1st Half',
    #                       frame > maxframe/2 ~ '2nd Half'),
         Time = case_when(frame <= maxframe/3 ~ '1',
                           between(frame, maxframe/3, maxframe*2/3)  ~  '2',
                          frame > maxframe*2/3 ~ '3'))

initialmomentumdata <- normalizedmomentumdata |> calcInitialMomentum()

meantotalmomentumdata  <- turnmomentumdata |> 
  group_by(ID, Time) |> 
  summarise(Individual = first(Individual),
            Speed.group = first(Speed.group),
            meanmomentum.gnormdegps = mean(momentum.gnormdegps),
            meanlinearmomentum.gnormcmps = mean(linearmomentum.gnormcmps)) |> 
  rbind(initialmomentumdata)
```
```{r}
meantotalmomentumdata
```

## Now we plot
```{r}
scaleFUN <- function(x) sprintf("%.1f", x)

# meantotalmomentumdata$Time = factor(meantotalmomentumdata$Time, levels=c("Pre", "1st Half", "2nd Half"))
meantotalmomentumdata$Time = factor(meantotalmomentumdata$Time, levels=c("Pre", "1", "2", "3"))
meantotalmomentumdata$Speed.group = factor(meantotalmomentumdata$Speed.group, levels=c("Slow", "Fast"))

meantotalmomentumdata |> angularmomentumplot('Slow','deepskyblue', -.1, 2)+ scale_y_continuous(labels=scaleFUN)
ggsave('AngularMomentumSlow.jpg', path = here("ProcessedData/Figures/"), width = 3.25, height = 2, dpi = 300)
meantotalmomentumdata |> angularmomentumplot('Fast','firebrick', -.1, 3.45) + scale_y_continuous(labels=scaleFUN)
ggsave('AngularMomentumFast.jpg', path = here("ProcessedData/Figures/"), width = 3.25, height = 2, dpi = 300)

meantotalmomentumdata |> linearmomentumplot('Slow', 'deepskyblue', 0, .03)+ scale_y_continuous(labels=scaleFUN)
ggsave('LinearMomentumSlow.jpg', path = here("ProcessedData/Figures/"), width = 3.25, height = 2, dpi = 300)
meantotalmomentumdata |> linearmomentumplot('Fast', 'firebrick', 0, .06)+ scale_y_continuous(labels=scaleFUN)
ggsave('LinearMomentumFast.jpg', path = here("ProcessedData/Figures/"), width = 3.25, height = 2, dpi = 300)
```

## Put together the momentum plots

```{r}
angmom.plot <- meantotalmomentumdata |> 
  group_by(ID) |> 
  ggplot(aes(x = Time, y = meanmomentum.gnormdegps, 
             color=Speed.group)) +
  geom_line(aes(group=ID),
            alpha=0.2) +
  stat_eye(mapping = aes(shape = Speed.group),
           point_interval = "mean_qi") +
  facet_grid(cols = vars(Speed.group)) +
  scale_shape_manual(values = c(3, 3)) +
  scale_color_manual(values = c('deepskyblue', 'firebrick')) +
  labs(y = 'Angular momentum\n(norm)')


linmom.plot <- meantotalmomentumdata |> 
  group_by(ID) |> 
  ggplot(aes(x = Time, y = meanlinearmomentum.gnormcmps, 
             color=Speed.group)) +
  geom_line(aes(group=ID),
            alpha=0.2) +
  stat_eye(mapping = aes(shape = Speed.group),
           point_interval = "mean_hdi") +
  facet_grid(cols = vars(Speed.group)) +
  scale_shape_manual(values = c(3, 3)) +
  scale_color_manual(values = c('deepskyblue', 'firebrick')) +
  labs(y = 'Linear momentum\n(norm)')

(angmom.plot / linmom.plot) & 
  (theme_classic() +
     theme(legend.position = 'none',
           strip.text.x = element_blank()))
```

## Set up to compare momentum

```{r}
meantotalmomentumwide <-
  meantotalmomentumdata |> 
  rename(linmom.norm=meanlinearmomentum.gnormcmps,
         angmom.norm=meanmomentum.gnormdegps) |> 
  group_by(ID) |> 
  mutate(is_mom_decrease = linmom.norm[Time == '1'] <
           linmom.norm[Time == 'Pre']) |> 
  arrange(ID, Time, Speed.group) |> 
  pivot_wider(names_from = "Time",
              values_from = c(linmom.norm, angmom.norm)) |> 
  mutate(linmom.d1p = linmom.norm_1 - linmom.norm_Pre,
         linmom.d21 = linmom.norm_2 - linmom.norm_1,
         linmom.d32 = linmom.norm_3 - linmom.norm_2,
         angmom.d1p = angmom.norm_1 - angmom.norm_Pre,
         angmom.d2p = angmom.norm_2 - angmom.norm_Pre,
         angmom.d21 = angmom.norm_2 - angmom.norm_1,
         angmom.d32 = angmom.norm_3 - angmom.norm_2)
```

```{r}
meantotalmomentumwide |> 
  ggduo(columnsX = c("linmom.norm_Pre",
                     "linmom.norm_1",
                      "linmom.d1p",
                      "linmom.d21"
                      ),
        columnsY = c("angmom.norm_1",
                     "angmom.d2p",
                     "angmom.d21",
                      "angmom.d32",
                     "linmom.d21",
                      "linmom.d32"
                      ),
          mapping = aes(color = Speed.group), #, shape = is_mom_decrease),
          progress=FALSE,
          legend = c(4, 2),
        types = list(continuous = wrap("smooth_lm"))) +
  geom_hline(yintercept=0) +
  geom_vline(xintercept=0)
  # scale_shape_manual(values = c(2, 19))
```

## Plot the tradeoff in linear and angular momentum

```{r}
tradeoff.plot <-
  meantotalmomentumwide |> 
  ggplot(aes(x = linmom.norm_Pre, y = angmom.d21, color = Speed.group)) +
  geom_point(aes(shape = Speed.group)) +
  geom_smooth(method = 'lm') +
  scale_color_manual(values = c('deepskyblue', 'firebrick')) +
  scale_shape_manual(values = c(2, 19)) +
  labs(x = 'Linear momentum (Pre)',
       y = 'Change in angular momentum\n(Time 2 - Time 1)') +
  theme_classic() +
  theme(legend.position = 'none')
```

And put the entire plot together

```{r}
((angmom.plot / linmom.plot) | tradeoff.plot) & 
  (theme_classic() +
     theme(legend.position = 'none',
           strip.text.x = element_blank()))

ggsave(here('ProcessedData','Figures','momentum-tradeoff.pdf'),
       width = 6.5, height = 3, units = 'in')
```


### Momentum stats

Make a long data frame so that we can easily run the same tests on linear and angular momentum

```{r}
meantotalmomentumlong <-
  meantotalmomentumdata |> 
  rename(angularmomentum.gnormdegps = meanmomentum.gnormdegps,
         linearmomentum.gnormcmps = meanlinearmomentum.gnormcmps) |> 
  ungroup() |> 
  pivot_longer(cols = c(angularmomentum.gnormdegps, linearmomentum.gnormcmps),
               values_to = "momentum",
               names_to = "type") |> 
  mutate(type = case_match(type, 
                           "angularmomentum.gnormdegps" ~ "angular",
                           "linearmomentum.gnormcmps" ~ "linear"))
```

This runs the comparison of momentum relative to time for slow vs fast turns and for linear and angular momentum.
```{r}
momentumstats <-
  meantotalmomentumlong |> 
  group_by(Speed.group, type) |> 
  nest(data = c(momentum,
                Time, Individual, ID)) |> 
  # make all of the stats models
  mutate(momentum.model = 
           purrr::map(data, 
                      \(df) lmer(momentum ~ Time + (1 | Individual),
                                 data = df)),
         momentum.anova = purrr::map(
           momentum.model, \(mod) broom::tidy(Anova(mod)))
  ) |> 
  # get the overall p values
  unnest(momentum.anova) |> 
  # pull out the contrasts
  mutate(contrasts = 
           purrr::map(momentum.model,
                      \(model) estimate_contrasts(model, 
                                                  contrast = "Time",
                                                  p_adjust = 'sidak'))
  ) |> 
  # and the compact letter displays for comparisons
  mutate(cld = 
           purrr::map(momentum.model,
                      \(model) cld(emmeans(model, specs = ~ Time))))

momentumstats
```


### Show the contrasts

```{r}
momentumstats |> 
  unnest(contrasts, names_sep = '.') |> 
  dplyr::rename(Speed = Speed.group) |> 
  dplyr::select(Speed, type, p.value, contains("Level"),
         contrasts.Difference, contrasts.p) |> 
  dplyr::rename(L1 = contrasts.Level1,
                L2 = contrasts.Level2,
                diff = contrasts.Difference) |> 
  mutate(sigcont = contrasts.p < 0.01) |> 
  relocate(sigcont, .after=L2)
```

### Show the CLDs

```{r}
momentumstats |> 
  unnest(cld, names_sep = '.') |> 
  rename(Speed = Speed.group,
         group = cld..group,
         Time = cld.Time) |> 
  dplyr::select(Speed, type, p.value, Time, group) |> 
  arrange(Speed, type, Time)

```

### Compare pre momentum

```{r}
meantotalmomentumlong |> 
  filter(Time == "Pre") |> 
  group_by(type) |> 
  nest(data = c(momentum, Speed.group,
                Time, Individual, ID)) |> 
  mutate(momentum.model = 
           purrr::map(data, 
                      \(df) lmer(momentum ~ Speed.group + (1 | Individual),
                                 data = df)),
         momentum.anova = purrr::map(
           momentum.model, \(mod) broom::tidy(Anova(mod)))
  ) |> 
  unnest(momentum.anova)
```

```{r}
meantotalmomentumwide |> 
  group_by(Speed.group) |> 
  dplyr::select(Individual, ID, Speed.group,
                linmom.norm_Pre, angmom.d21) |> 
  nest(data = c(Individual, ID, linmom.norm_Pre, angmom.d21)) |> 
  mutate(regr.model = 
           purrr::map(data, 
                      \(df) lmer(angmom.d21 ~ linmom.norm_Pre + (1 | Individual),
                                 data = df)),
         momentum.anova = purrr::map(
           regr.model, \(mod) broom::tidy(Anova(mod)))
  ) |> 
  unnest(momentum.anova)
```

